### 2018-09-23 vmmsh-2018/blg.tex
[vmmsh] Bibiliography- WIP

\begin{thebibliography}{99}
\bibitem{erdos1945integral}
\textsl{Erdös} \textsl{P.} Integral distances // Bulletin of the
American Mathematical Society. — 1945. — Т. 51, No 12. — С. 996.
\bibitem{anning1945integral}
\textsl{Anning} \textsl{N. H.}, \textsl{Erdös} \textsl{P.}
Integral distances // Bulletin of the American Mathematical
Society. — 1945. — Т. 51, No 8. — С. 598—600.
\bibitem{solymosi2003note}
\textsl{Solymosi} \textsl{J.} Note on integral distances //
Discrete \& Computational Geometry. — 2003. — Т. 30, No 2. —
С. 337—342.
\bibitem{anning1915discussions}
\textsl{Anning} \textsl{N.} Discussions: Relating to a Geometric
Representation of Integral Solutions of Certain Quadratic
Equations // The American Mathematical Monthly. — 1915. —
Т. 22, No 9. — С. 321—321.
\bibitem{harborth1993upper}
\textsl{Harborth} \textsl{H.},
\textsl{Möller} \textsl{M.} An
diameter of integral point sets
Geometry. — 1993. — Т. 9, No 4.
\textsl{Kemnitz} \textsl{A.},
upper bound for the minimum
// Discrete \& Computational
— С. 427—432.\bibitem{piepmeyer1996maximum}
\textsl{Piepmeyer} \textsl{L.} The maximum number of odd
integral distances between points in the plane // Discrete \&
Computational Geometry. — 1996. — Т. 16, No 1. — С. 113—115.
\bibitem{kurz2008bounds}
\textsl{Kurz} \textsl{S.}, \textsl{Laue} \textsl{R.} Bounds for
the minimum diameter of integral point sets // arXiv preprint
arXiv:0804.1296. — 2008.
\bibitem{our-vvmsh-2018}
\textsl{Авдеев} \textsl{Н. Н.} Некоторые примеры систем
Эрдёша // Современные методы теории краевых задач :
материалы международной конференции «Понтрягинские
чтения — XXIX», посвященной 90-летию Владимира
Александровича Ильина (2–6 мая 2018 г.) — 2018. — С. 30—31.
\bibitem{locatelli2002packing}
\textsl{Locatelli} \textsl{M.}, \textsl{Raber} \textsl{U.} Packing
equal circles in a square: a deterministic global optimization
approach // Discrete Applied Mathematics. — 2002. — Т. 122,
No 1—3. — С. 139—166.
\bibitem{costa2013valid}
\textsl{Costa} \textsl{A.} Valid constraints for the point packing
in a square problem // Discrete Applied Mathematics. — 2013. —
Т. 161, No 18. — С. 2901—2909.
\bibitem{markot2005newverified}
\textsl{Markót} \textsl{M. C.}, \textsl{Csendes} \textsl{T.} A
new verified optimization technique for the "packing circles in
a unit square" problems // SIAM Journal on Optimization. —
2005. — Т. 16, No 1. — С. 193—219.
\bibitem{degroot1990optimal10points}
\textsl{DeGroot} \textsl{C.}, \textsl{Peikert} \textsl{R.},
\textsl{Würtz} \textsl{D.} The optimal packing of ten equal
circles in a square. — Eidgenössische Technische Hochschule
Zürich, Interdisziplinäres Projektzentrum für Supercomputing,
1990.\bibitem{kemnitz1988punktmengen}
\textsl{Kemnitz} \textsl{A.} Punktmengen mit ganzzahligen
Abständen. — 1988.
\bibitem{Bukhstab-number-theory}
\textsl{Бухштаб} \textsl{А.} Теория чисел. — 1966. — С. 384.
\bibitem{our-mkmitu-2016}
\textsl{Авдеев} \textsl{Н. Н.} Об отыскании множеств точек на
плоскости с целочисленными расстояниями // Математическое
и компьютерное моделирование, информационные технологии
управления: сб. тр. Школы для студентов, аспирантов и
молодых ученых «МКМИТУ-2016». — 2016. — С. 15—19.
\bibitem{our-ped-2017}
\textsl{Авдеев} \textsl{Н. Н.} О множествах точек на
плоскости с целочисленными расстояниями и алгоритмах их
отыскания // Некоторые вопросы анализа, алгебры, геометрии
и математического образования: материалы международной
молодежной научной школы «Актуальные направления
математического анализа и смежные вопросы». — 2017. — Т. 7,
No I. — С. 9—10.
\bibitem{polygons-on-lattices}
\textsl{Вавилов} \textsl{В.}, \textsl{Устинов}
Многоугольники на решетках. — 2006.
\textsl{А.}
\bibitem{Kohnert2006IntegralPS}
\textsl{Kohnert} \textsl{A.}, \textsl{Kurz} \textsl{S.} Integral
point sets over $Z^m_n$
// Discrete Applied Mathematics. — 2006. —
Т. 157 (2009). — С. 2105—2117.
\bibitem{kurz2005characteristic}
\textsl{Kurz} \textsl{S.} On the characteristic of integral point
sets in // arXiv preprint math/0511704. — 2005.
\bibitem{harborth1985diameters}
\textsl{Harborth} \textsl{H.}, \textsl{Kemnitz} \textsl{A.}
Diameters of integral point sets // Colloquia Mathematica
Societatis Jano Bolyai. Т. 48. — 1985.
\bibitem{kanold1981punktmengen}
\textsl{Kanold}
\textsl{H.-J.}
Über
Punktmengen
im
$k$-dimensionalen euklidischen Raum // Abh. Braunschw. Wiss.
Ges. — 1981. — Т. 32, No 55—65. — С. 252.
\bibitem{nozaki2013lower}
\textsl{Nozaki} \textsl{H.} Lower bounds for the minimum
diameter of integral point sets // Australasian Journal of
Combinatorics. — 2013. — Т. 56. — С. 139—143.
\bibitem{piepmeyer1988raumliche}
\textsl{Piepmeyer}
\textsl{L.}
Räumliche
ganzzahlige
Punktmengen: дис. . . . канд. / Piepmeyer L. — Master’s
thesis, TU Braunschweig, 1988.
\bibitem{harborth1991point}
\textsl{Harborth} \textsl{H.}, \textsl{Piepmeyer} \textsl{L.}
Point sets with small integral distances // Applied geometry and
discrete mathematics: the Victor Klee Festschrift. Т. 4. — American
Mathematical Soc., 1991.
\bibitem{harborth1998integral}
\textsl{Harborth} \textsl{H.} Integral distances in point sets //
Karl der Grosse und sein Nachwirken. 1200 Jahre Kultur und
Wissenschaft in Europa: Band II, Mathematisches Wissen. —
1998. — С. 213—224.
\bibitem{kurz2006konstruktion}
\textsl{Kurz} \textsl{S.} Konstruktion und Eigenschaften
ganzzahliger Punktmengen: дис. . . . канд. / Kurz Sascha. — 2006.
\bibitem{kurz2008minimum}
\textsl{Kurz} \textsl{S.}, \textsl{Wassermann} \textsl{A.} On
the minimum diameter of plane integral point sets // arXiv
preprint arXiv:0804.1307. — 2008.
\bibitem{our-mz-rus}
\textsl{Авдеев} \textsl{Н. Н.}, \textsl{Семёнов} \textsl{Е. М.}
О множествах точек на плоскости с целочисленными
расстояниями // Математические заметки. — 2016. — Т. 100,
No 5. — С. 757—761.\bibitem{kreisel2008heptagon}
\textsl{Kreisel} \textsl{T.}, \textsl{Kurz} \textsl{S.} There are
integral heptagons, no three points on a line, no four on a circle //
Discrete \& Computational Geometry. — 2008. — Т. 39, No 4. —
С. 786—790.
\bibitem{brass2006research}
\textsl{Brass} \textsl{P.}, \textsl{Moser} \textsl{W. O.},
\textsl{Pach} \textsl{J.} Research problems in discrete
geometry. — Springer Science \& Business Media, 2006.
\bibitem{guy2013unsolved}
\textsl{Guy} \textsl{R.} Unsolved problems in number theory. Т.
1. — Springer Science \& Business Media, 2013.
\bibitem{noll1989nclusters}
\textsl{Noll} \textsl{L. C.}, \textsl{Bell} \textsl{D. I.} n-clusters
for $1 < n < 7$ // Mathematics of Computation. — 1989. — Т. 53,
No 187. — С. 439—444.
\bibitem{kurz2013constructing}
Constructing 7-clusters / S. Kurz [и др.] // arXiv preprint
arXiv:1312.2318. — 2013.
\bibitem{graham1974there}
\textsl{Graham} \textsl{R.}, \textsl{Rothschild} \textsl{B.},
\textsl{Straus} \textsl{E.} Are there $n + 2$ points in $E^n$ with odd
integral distances? // The American Mathematical Monthly. —
1974. — Т. 81, No 1. — С. 21—25.
\bibitem{ulam1960collection}
\textsl{Ulam} \textsl{S. M.} A collection of mathematical
problems. Т. 8. — Interscience Publishers, 1960.
\bibitem{garibaldi2005erdos}
\textsl{Garibaldi} \textsl{J.}, \textsl{Iosevich} \textsl{A.} The
Erdos distance problem: lecture notes. — 2005.
\bibitem{garibaldi2011erdos}
\textsl{Garibaldi} \textsl{J.}, \textsl{Iosevich} \textsl{A.},
\textsl{Senger} \textsl{S.} The Erdös Distance Problem. Т. 56. —
American Mathematical Soc., 2011.\bibitem{erdos1946sets}
\textsl{Erdös} \textsl{P.} On sets of distances of n points //
The American Mathematical Monthly. — 1946. — Т. 53, No 5. —
С. 248—250.
\bibitem{guth2015erdos}
\textsl{Guth} \textsl{L.}, \textsl{Katz} \textsl{N. H.} On the
Erdos distinct distances problem in the plane // Annals of
Mathematics. — 2015. — Т. 181. — С. 155—190.
\end{thebibliography}