### 2018-09-23 vmmsh-2018/blg.cp1251.tex

\begin{thebibliography}{99}
\bibitem{erdos1945integral}
\textsl{Erd\"{o}s} \textsl{P.} Integral distances /\!/ Bulletin of the
American Mathematical Society. --- 1945. --- Т. 51, No 12. --- С. 996.
\bibitem{anning1945integral}
\textsl{Anning} \textsl{N. H.}, \textsl{Erd\"{o}s} \textsl{P.}
Integral distances /\!/ Bulletin of the American Mathematical
Society. --- 1945. --- Т. 51, No 8. --- С. 598---600.
\bibitem{solymosi2003note}
\textsl{Solymosi} \textsl{J.} Note on integral distances /\!/
Discrete \& Computational Geometry. --- 2003. --- Т. 30, No 2. ---
С. 337---342.
\bibitem{anning1915discussions}
\textsl{Anning} \textsl{N.} Discussions: Relating to a Geometric
Representation of Integral Solutions of Certain Quadratic
Equations /\!/ The American Mathematical Monthly. --- 1915. ---
Т. 22, No 9. --- С. 321---321.
\bibitem{harborth1993upper}
\textsl{Harborth} \textsl{H.},
\textsl{M\"{o}ller} \textsl{M.} An
diameter of integral point sets
Geometry. --- 1993. --- Т. 9, No 4.
\textsl{Kemnitz} \textsl{A.},
upper bound for the minimum
/\!/ Discrete \& Computational Geometry
--- С. 427---432.\bibitem{piepmeyer1996maximum}
\textsl{Piepmeyer} \textsl{L.} The maximum number of odd
integral distances between points in the plane /\!/ Discrete \&
Computational Geometry. --- 1996. --- Т. 16, No 1. --- С. 113---115.
\bibitem{kurz2008bounds}
\textsl{Kurz} \textsl{S.}, \textsl{Laue} \textsl{R.} Bounds for
the minimum diameter of integral point sets /\!/ arXiv preprint
arXiv:0804.1296. --- 2008.
\bibitem{our-vvmsh-2018}
\textsl{Авдеев} \textsl{Н. Н.} Некоторые примеры систем
Эрдёша /\!/ Современные методы теории краевых задач :
материалы международной конференции <<Понтрягинские
чтения --- XXIX>>, посвященной 90-летию Владимира
Александровича Ильина (2--6 мая 2018 г.) --- 2018. --- С. 30---31.
\bibitem{locatelli2002packing}
\textsl{Locatelli} \textsl{M.}, \textsl{Raber} \textsl{U.} Packing
equal circles in a square: a deterministic global optimization
approach /\!/ Discrete Applied Mathematics. --- 2002. --- Т. 122,
No 1---3. --- С. 139---166.
\bibitem{costa2013valid}
\textsl{Costa} \textsl{A.} Valid constraints for the point packing
in a square problem /\!/ Discrete Applied Mathematics. --- 2013. ---
Т. 161, No 18. --- С. 2901---2909.
\bibitem{markot2005newverified}
\textsl{Mark\'{o}t} \textsl{M. C.}, \textsl{Csendes} \textsl{T.} A
new verified optimization technique for the "packing circles in
a unit square" problems /\!/ SIAM Journal on Optimization. ---
2005. --- Т. 16, No 1. --- С. 193---219.
\bibitem{degroot1990optimal10points}
\textsl{DeGroot} \textsl{C.}, \textsl{Peikert} \textsl{R.},
\textsl{W\"{u}rtz} \textsl{D.} The optimal packing of ten equal
circles in a square. --- Eidgen\"{o}ssische Technische Hochschule
Z\"{u}rich, Interdisziplin\"{a}res Projektzentrum f\"{u}r Supercomputing,
1990.\bibitem{kemnitz1988punktmengen}
\textsl{Kemnitz} \textsl{A.} Punktmengen mit ganzzahligen
Abst\"{a}nden. --- 1988.
\bibitem{Bukhstab-number-theory}
\textsl{Бухштаб} \textsl{А.} Теория чисел. --- 1966. --- С. 384.
\bibitem{our-mkmitu-2016}
\textsl{Авдеев} \textsl{Н. Н.} Об отыскании множеств точек на
плоскости с целочисленными расстояниями /\!/ Математическое
и компьютерное моделирование, информационные технологии
управления: сб. тр. Школы для студентов, аспирантов и
молодых ученых <<МКМИТУ-2016>>. --- 2016. --- С. 15---19.
\bibitem{our-ped-2017}
\textsl{Авдеев} \textsl{Н. Н.} О множествах точек на
плоскости с целочисленными расстояниями и алгоритмах их
отыскания /\!/ Некоторые вопросы анализа, алгебры, геометрии
и математического образования: материалы международной
молодежной научной школы <<Актуальные направления
математического анализа и смежные вопросы>>. --- 2017. --- Т. 7,
No I. --- С. 9---10.
\bibitem{polygons-on-lattices}
\textsl{Вавилов} \textsl{В.}, \textsl{Устинов}
Многоугольники на решетках. --- 2006.
\textsl{А.}
\bibitem{Kohnert2006IntegralPS}
\textsl{Kohnert} \textsl{A.}, \textsl{Kurz} \textsl{S.} Integral
point sets over $Z^m_n$
/\!/ Discrete Applied Mathematics. --- 2006. ---
Т. 157 (2009). --- С. 2105---2117.
\bibitem{kurz2005characteristic}
\textsl{Kurz} \textsl{S.} On the characteristic of integral point
sets in /\!/ arXiv preprint math/0511704. --- 2005.
\bibitem{harborth1985diameters}
\textsl{Harborth} \textsl{H.}, \textsl{Kemnitz} \textsl{A.}
Diameters of integral point sets /\!/ Colloquia Mathematica
Societatis Jano Bolyai. Т. 48. --- 1985.
\bibitem{kanold1981punktmengen}
\textsl{Kanold}
\textsl{H.-J.}
\"{U}ber
Punktmengen
im
$k$-dimensionalen euklidischen Raum /\!/ Abh. Braunschw. Wiss.
Ges. --- 1981. --- Т. 32, No 55---65. --- С. 252.
\bibitem{nozaki2013lower}
\textsl{Nozaki} \textsl{H.} Lower bounds for the minimum
diameter of integral point sets /\!/ Australasian Journal of
Combinatorics. --- 2013. --- Т. 56. --- С. 139---143.
\bibitem{piepmeyer1988raumliche}
\textsl{Piepmeyer}
\textsl{L.}
R\"{a}umliche
ganzzahlige
Punktmengen: дис. . . . канд. / Piepmeyer L. --- Master’s
thesis, TU Braunschweig, 1988.
\bibitem{harborth1991point}
\textsl{Harborth} \textsl{H.}, \textsl{Piepmeyer} \textsl{L.}
Point sets with small integral distances /\!/ Applied geometry and
discrete mathematics: the Victor Klee Festschrift. Т. 4. --- American
Mathematical Soc., 1991.
\bibitem{harborth1998integral}
\textsl{Harborth} \textsl{H.} Integral distances in point sets /\!/
Karl der Grosse und sein Nachwirken. 1200 Jahre Kultur und
Wissenschaft in Europa: Band II, Mathematisches Wissen. ---
1998. --- С. 213---224.
\bibitem{kurz2006konstruktion}
\textsl{Kurz} \textsl{S.} Konstruktion und Eigenschaften
ganzzahliger Punktmengen: дис. . . . канд. / Kurz Sascha. --- 2006.
\bibitem{kurz2008minimum}
\textsl{Kurz} \textsl{S.}, \textsl{Wassermann} \textsl{A.} On
the minimum diameter of plane integral point sets /\!/ arXiv
preprint arXiv:0804.1307. --- 2008.
\bibitem{our-mz-rus}
\textsl{Авдеев} \textsl{Н. Н.}, \textsl{Семёнов} \textsl{Е. М.}
О множествах точек на плоскости с целочисленными
расстояниями /\!/ Математические заметки. --- 2016. --- Т. 100,
No 5. --- С. 757---761.\bibitem{kreisel2008heptagon}
\textsl{Kreisel} \textsl{T.}, \textsl{Kurz} \textsl{S.} There are
integral heptagons, no three points on a line, no four on a circle /\!/
Discrete \& Computational Geometry. --- 2008. --- Т. 39, No 4. ---
С. 786---790.
\bibitem{brass2006research}
\textsl{Brass} \textsl{P.}, \textsl{Moser} \textsl{W. O.},
\textsl{Pach} \textsl{J.} Research problems in discrete
geometry. --- Springer Science \& Business Media, 2006.
\bibitem{guy2013unsolved}
\textsl{Guy} \textsl{R.} Unsolved problems in number theory. Т.
1. --- Springer Science \& Business Media, 2013.
\bibitem{noll1989nclusters}
\textsl{Noll} \textsl{L. C.}, \textsl{Bell} \textsl{D. I.} n-clusters
for $1 < n < 7$ /\!/ Mathematics of Computation. --- 1989. --- Т. 53,
No 187. --- С. 439---444.
\bibitem{kurz2013constructing}
Constructing 7-clusters / S. Kurz [и др.] /\!/ arXiv preprint
arXiv:1312.2318. --- 2013.
\bibitem{graham1974there}
\textsl{Graham} \textsl{R.}, \textsl{Rothschild} \textsl{B.},
\textsl{Straus} \textsl{E.} Are there $n + 2$ points in $E^n$ with odd
integral distances? /\!/ The American Mathematical Monthly. ---
1974. --- Т. 81, No 1. --- С. 21---25.
\bibitem{ulam1960collection}
\textsl{Ulam} \textsl{S. M.} A collection of mathematical
problems. Т. 8. --- Interscience Publishers, 1960.
\bibitem{garibaldi2005erdos}
\textsl{Garibaldi} \textsl{J.}, \textsl{Iosevich} \textsl{A.} The
Erdos distance problem: lecture notes. --- 2005.
\bibitem{garibaldi2011erdos}
\textsl{Garibaldi} \textsl{J.}, \textsl{Iosevich} \textsl{A.},
\textsl{Senger} \textsl{S.} The Erd\"{o}s Distance Problem. Т. 56. ---
American Mathematical Soc., 2011.\bibitem{erdos1946sets}
\textsl{Erd\"{o}s} \textsl{P.} On sets of distances of n points /\!/
The American Mathematical Monthly. --- 1946. --- Т. 53, No 5. ---
С. 248---250.
\bibitem{guth2015erdos}
\textsl{Guth} \textsl{L.}, \textsl{Katz} \textsl{N. H.} On the
Erdos distinct distances problem in the plane /\!/ Annals of
Mathematics. --- 2015. --- Т. 181. --- С. 155---190.
\end{thebibliography}