### 2018-09-23 vmmsh-2018/blg.cp1251.tex.tex

\begin{thebibliography}{99}
\bibitem{erdos1945integral}
\textsl{Erd\"{o}s} \textsl{P.} Integral distances /\!/ Bulletin of the
American Mathematical Society. Ч 1945. Ч “. 51, No 12. Ч —. 996.
\bibitem{anning1945integral}
\textsl{Anning} \textsl{N. H.}, \textsl{Erd\"{o}s} \textsl{P.}
Integral distances /\!/ Bulletin of the American Mathematical
Society. Ч 1945. Ч “. 51, No 8. Ч —. 598Ч600.
\bibitem{solymosi2003note}
\textsl{Solymosi} \textsl{J.} Note on integral distances /\!/
Discrete \& Computational Geometry. Ч 2003. Ч “. 30, No 2. Ч
—. 337Ч342.
\bibitem{anning1915discussions}
\textsl{Anning} \textsl{N.} Discussions: Relating to a Geometric
Representation of Integral Solutions of Certain Quadratic
Equations /\!/ The American Mathematical Monthly. Ч 1915. Ч
“. 22, No 9. Ч —. 321Ч321.
\bibitem{harborth1993upper}
\textsl{Harborth} \textsl{H.},
\textsl{M\"{o}ller} \textsl{M.} An
diameter of integral point sets
Geometry. Ч 1993. Ч “. 9, No 4.
\textsl{Kemnitz} \textsl{A.},
upper bound for the minimum
/\!/ Discrete \& Computational Geometry
Ч —. 427Ч432.\bibitem{piepmeyer1996maximum}
\textsl{Piepmeyer} \textsl{L.} The maximum number of odd
integral distances between points in the plane /\!/ Discrete \&
Computational Geometry. Ч 1996. Ч “. 16, No 1. Ч —. 113Ч115.
\bibitem{kurz2008bounds}
\textsl{Kurz} \textsl{S.}, \textsl{Laue} \textsl{R.} Bounds for
the minimum diameter of integral point sets /\!/ arXiv preprint
arXiv:0804.1296. Ч 2008.
\bibitem{our-vvmsh-2018}
\textsl{јвдеев} \textsl{Ќ. Ќ.} Ќекоторые примеры систем
ЁрдЄша /\!/ —овременные методы теории краевых задач :
материалы международной конференции <<ѕонтр€гинские
чтени€ Ч XXIX>>, посв€щенной 90-летию ¬ладимира
јлександровича »льина (2Ц6 ма€ 2018 г.) Ч 2018. Ч —. 30Ч31.
\bibitem{locatelli2002packing}
\textsl{Locatelli} \textsl{M.}, \textsl{Raber} \textsl{U.} Packing
equal circles in a square: a deterministic global optimization
approach /\!/ Discrete Applied Mathematics. Ч 2002. Ч “. 122,
No 1Ч3. Ч —. 139Ч166.
\bibitem{costa2013valid}
\textsl{Costa} \textsl{A.} Valid constraints for the point packing
in a square problem /\!/ Discrete Applied Mathematics. Ч 2013. Ч
“. 161, No 18. Ч —. 2901Ч2909.
\bibitem{markot2005newverified}
\textsl{Mark