### 2019-05-08 ped2018/AvdeevNN_3_geom.tex
Theses for PED2018

\documentclass[11pt]{article}
\usepackage[left=20mm, right=20mm, top=20mm, bottom=20mm, includefoot, a5paper]{geometry}
\usepackage[cp1251]{inputenc}
\usepackage[english,russian]{babel} %ѕоправлено. ѕозволит нормально переносить слова в английском тексте.
\usepackage[intlimits]{amsmath}
\interdisplaylinepenalty=2500
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{latexsym}
\pagestyle{empty}

\usepackage{graphicx} % ј как иначе вставл€ть картинки??

\begin{document}

\addcontentsline{toc}{section}{
	Avdeev N.\,N.
	On integral point sets in special position
}

\small UDC 511.95 + 514.112
\begin{center}
\bf
ON INTEGRAL POINT SETS IN SPECIAL POSITION%
\footnote{This work was supported by the Russian Science
Foundation under grant 16-11-10125.}\\
\small Avdeev N.\,N.\\
\small Voronezh State University\\
\small avdeev@math.vsu.ru, nickkolok@mail.ru\\
\end{center}

% ќбнул€ю счЄтчик на вс€кий случай
\setcounter{equation}{0}

% Switching to English
\selectlanguage{english}
\setcounter{figure}{0}

\textbf{Definition 1.}
A planar integral point set (PIPS) is a set $M$ of noncollinear points in plane $\mathbb{R}^2$
such that for any pair of points $M_1, M_2\in M$ the distance between $M_1$ and $M_2$ is integral.
Every PIPS is finite [1].

Some general constructions of PIPS are known,
e.g. a concyclic PIPS of an arbitrary power [2]
and a PIPS of an arbitrary power $n$ such that $(n-1)$ points are collinear [3].
However, there is no known algorithm to construct a PIPS of arbitrary power $n$ which is contained in 2 parallel straight lines,
no more than $n-2$ points on each line.

Examples of such PIPS for $n=7,8,9$ can be found in [4,5].
Below an example for $n=10$ is given.
We use notation
\begin{equation*}
	\sqrt{p}/q * \{ (x_1,y_1), ..., (x_n, y_n)  \}
	:=
	\left\{ \left(\frac{x_1}{q},\frac{y_1\sqrt{p}}{q}\right), ..., \left(\frac{x_n}{q},\frac{y_n\sqrt{p}}{q}\right)  \right\}
	.
\end{equation*}
\\
Figure~\ref{Avdeev_10_158_1538485325776.crop.eps}: $
\sqrt{70}/79 *
\{
( \pm6241 , 0),
( -873 , -528),
( 2093 , 408),
( 4838 , 138),
$\\$
( -3275 , 936),
( -1872 , 798),
( 5265 , 96),
( -2299 , 840),
( 873 , 528)
\}
$.
%\\
%Figure~\ref{Avdeev_10_222_1538488624864.crop.eps}
%
%$
%\sqrt{70}/37 *
%\{
%( -4107 , 0),
%( 4107 , 0),
%( 840 , 198)
%$\\$
%( -2622 , -90),
%( -1599 , -152),
%( -246 , -234),
%$\\$
%( 381 , -272),
%( 1206 , -322),
%( 1833 , -360),
%( 3186 , -442),
%\}
%$

\begin{figure}[h]
\center
	\includegraphics[width=0.7\linewidth]{Avdeev_10_158_1538485325776.eps}
	\caption{PIPS of power 10 and diameter 158}\label{Avdeev_10_158_1538485325776.crop.eps}%-converted-to-crop.ps}
\end{figure}

%\begin{figure}[h!]
%	\center
%	\includegraphics[width=0.7\linewidth]{Avdeev_10_222_1538488624864.eps}
%	\caption{PIPS of power 10 and diameter 222}\label{Avdeev_10_222_1538488624864.crop.eps}%-converted-to-crop.ps}
%\end{figure}

 \centerline{\bf References}

1.
Erd\"{o}s P. Integral distances //Bulletin of the American Mathematical Society. Ц 1945. Ц V. 51. Ц No. 12. Ц P. 996.

2.
Harborth H., Kemnitz A., M\"{o}ller M. An upper bound for the minimum diameter of integral point sets
//Discrete \& Computational Geometry. Ц 1993. Ц V. 9. Ц No. 4. Ц P. 427-432.

3.
Solymosi J. Note on integral distances //Discrete \& Computational Geometry. Ц 2003. Ц V. 30. Ц No. 2. Ц P. 337-342.

\selectlanguage{russian}
4.
јвдеев Ќ.\,Ќ.
ќб отыскании целоудалЄнных множеств специального вида
//
—б. тр. ћеждународной конференции Ђјктуальные проблемы прикладной математики, информатики и механикиї
Ц ¬оронеж, 2018. Ц в печати. (in Russian)

\selectlanguage{english}

5.
Guy~R. Unsolved problems in number theory. V.
	1. --- Springer Science \& Business Media, 2013.

{\bf Avdeev Nikolai Nikolaevich},
master student of department of functional analysis and operator equations, Voronezh State University.
%магистрант каф. функционального анализа и операторных уравнений математического факультета ‘√Ѕќ” ¬ќ <<¬√”>>

% ќбнул€ю счЄтчик за собой
\setcounter{equation}{0}
\setcounter{figure}{0}

\end{document}