### 2020-01-03 pps-linear-bound/J220-AJC-review-reply-1.tex
[J220] Reply to the reviews

\documentclass[a4paper,14pt]{article} %размер бумаги устанавливаем А4, шрифт 12пунктов
\usepackage[T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[russian,english]{babel} %используем русский и английский языки с переносами
\usepackage{amssymb,amsfonts,amsmath,mathtext,enumerate,float,amsthm} %подключаем нужные пакеты расширений
\usepackage[unicode,colorlinks=true,citecolor=black,linkcolor=black]{hyperref}
%\usepackage[pdftex,unicode,colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{indentfirst} % включить отступ у первого абзаца
\usepackage[dvips]{graphicx} %хотим вставлять рисунки?
\graphicspath{{illustr/}}%путь к рисункам

\makeatletter
\renewcommand{\@biblabel}[1]{#1.} % Заменяем библиографию с квадратных скобок на точку:
\makeatother %Смысл этих трёх строчек мне непонятен, но поверим "Запискам дебианщика"

\usepackage{geometry} % Меняем поля страницы.
\geometry{left=2cm}% левое поле
\geometry{right=1cm}% правое поле
\geometry{top=2cm}% верхнее поле
\geometry{bottom=2cm}% нижнее поле

%\renewcommand{\theenumi}{\arabic{enumi}}% Меняем везде перечисления на цифра.цифра
%\renewcommand{\labelenumi}{\arabic{enumi}}% Меняем везде перечисления на цифра.цифра
%\renewcommand{\theenumii}{.\arabic{enumii}}% Меняем везде перечисления на цифра.цифра
%\renewcommand{\labelenumii}{\arabic{enumi}.\arabic{enumii}.}% Меняем везде перечисления на цифра.цифра
%\renewcommand{\theenumiii}{.\arabic{enumiii}}% Меняем везде перечисления на цифра.цифра
%\renewcommand{\labelenumiii}{\arabic{enumi}.\arabic{enumii}.\arabic{enumiii}.}% Меняем везде перечисления на цифра.цифра

\sloppy

\renewcommand\normalsize{\fontsize{14}{25.2pt}\selectfont}

%\usepackage[backend=biber,style=gost-numeric,sorting=nyt]{biblatex}
%\addbibresource{../common/notmy.bib}
%\addbibresource{../common/my.bib}

\begin{document}

Dear Editor and Referees,

Thank you for your attention paid to the paper.
I've done my best to cover all the remarks and advices,
and the present document is a detaled description of the improvements done.

First of all, I'd like to say special thanks to Referee 2 for the advice
to better guide the reader through the long list of technical auxiliary results.
As it was recommended, I've performed a major revision:
the structure of the paper has been rearranged.
Section 2 has been splitted into 3 smaller sections;
moreover, each section of main part has been divided into subsections
(see last paragraphs of Introduction for detailed description).

Due to such a rearrangement, numbers of most Lemmas, Theorems etc. have been changed,
and many results have been permutted to fit the new structure.
Thus, it is rather impossible to mention the page and line numbers of every change that has been made,
as it had been requested by the Editor.
As a possible solution, I've published all the workflow at my GitHub:
\\
https://github.com/nickkolok/sct-statya/commits/master
\\
As far as I put all my work on the topic into the single repository,
I've marked the commits concerning this paper with ``[J220]''.

As for Review 1, I agree with all the remarks completely.
\begin{enumerate}
	\item
		> Lemma 2.4: the reference is Russian, you may write a proof of Lemma 2.4 here in English

		Done. Lemma 2.4 is now Lemma 3.4 and has a self-contained proof.

	\item

		> page 10, line 3: $c_j \rightarrow 2c_J$ or you may eliminate this line.

		I've preferred the first option, i.e. replacement $c_j \rightarrow 2c_J$
\end{enumerate}

As for Review 2, I agree with all the remarks completely, except one.

Medium remarks:

\begin{enumerate}
	\item
		> The statement that integral point sets have to be finite is only true if we assume a dimension of at least 2.

		Fixed. Page 1, line -7.

	\item

		> As the author’s recent preprint
		``On diameter bounds for planar integral point sets in semi-general position''
		is not very long and overlaps the content of the present paper,
		it might be a good idea to summarize both to a single paper.

		This this the only remark for which I have to write a rebuttal for,
		and the reason is pretty simple:
		the paper ``On diameter bounds for planar integral point sets in semi-general position''
		has already been sent to another journal.
		Thus, it seems to be impossible to include its content into the present paper.
		Moreover, if the referees insist, I can
		\textbf{
		try to split the present paper into two ones:
		}
		one about the lower bound and one about integral point sets with distance 1.
		The present paper looks rather long, and the requested proofs (see below)
		have made it even more longer.
		However, if the paper will be published as a single one, then I'll also be very glad.

	\item

		> Please translate the titles of the non-english reference, like [8] and [3],
		into english or provide a link where those papers can be accessed.

		The translation has been done.
		However, the numbers of references have been changed, because
		the Editor requested to put the references in alphabetical order of authors.

	\item

		> Please add a self-contained proof for Lemma 2.4 since not all readers, including me, can read the one in [3].

		Done. Lemma 2.4 is now Lemma 3.4 and has a self-contained proof.

	\item

		> Is Lemma 2.6 also from [3] or have an obvious proof? Same for Lemma 2.7.

		Lemma 2.6 is now Lemma 4.2 and has a self-contained proof.
		\\
		Lemma 2.7 is now Lemma 4.4 and has a self-contained proof.

	\item

		> Was the inequality $d(2, 123) \leq d(2, 21491)$ used in the proof of Proposition 2.8?

		Yes, it was.
		The comment has been added before the proposition (which is now Proposition 3.5).

	\item

		> Please add a proof for Proposition 2.12.

		Thank you ver much for this remark.
		I have to admit that the result is not formulated correctly.
		The edits affect Definition 2.2 (ex Defintion 2.11), Proposition 2.3 (ex Proposition 2.12) and Lemma 2.4 (ex Lemma 2.13).
		Proposition 2.3 now has a self-contained proof.

	\item

		> The “blow up” procedure needs an explanation – not everybody is an expert on this.

		Construction 5.10 now provides the explanation.

\end{enumerate}

Minor remarks have been implemented precisely as they had been written.
As for minor remark 3, a brief proof for the corresponding Proposition 3.6 has been added.

%\printbibliography

\end{document}