### 2021-10-31 semi-general-suplinear-bound/blg-GOST.tex
[CHEB] Now the paper is accepted, and we can mention the author
Revert "We should not mention the author due to referring policy"

This reverts commit e2fa42531fa8f4e2330024a62bf335bfd5d8a0ef.

\begin{thebibliography}{99}
\bibitem{anning1945integral}
Anning N. H., Erdös P. Integral distances // Bulletin of the American Mathematical
Society. — 1945. — т. 51, No 8. — с. 598—600.
\bibitem{erdos1945integral}
Erdös P. Integral distances // Bulletin of the American Mathematical Society. — 1945. —
т. 51, No 12. — с. 996.
\bibitem{our-vmmsh-2018}
Авдеев Н. Н., Семёнов Е. М. Множества точек с целочисленными расстояниями на
плоскости и в евклидовом пространстве // Математический форум (Итоги науки. Юг
России). — 2018. — с. 217—236.
\bibitem{kurz2008bounds}
Kurz S., Laue R. Bounds for the minimum diameter of integral point sets // Australasian
Journal of Combinatorics. — 2007. — т. 39. — с. 233—240. — arXiv: 0804.1296.
\bibitem{kurz2008minimum}
Kurz S., Wassermann A. On the minimum diameter of plane integral point sets // Ars
Combinatoria. — 2011. — т. 101. — с. 265—287. — arXiv: 0804.1307.
\bibitem{antonov2008maximal}
Antonov A. R., Kurz S. Maximal integral point sets over $\mathbb{Z}^2$ // International Journal of
Computer Mathematics. — 2008. — т. 87, No 12. — с. 2653—2676. — arXiv: 0804.1280.
\bibitem{huff1948diophantine}
Huff G. B. Diophantine problems in geometry and elliptic ternary forms // Duke
Mathematical Journal. — 1948. — т. 15, No 2. — с. 443—453.
\bibitem{harborth1993upper}
Harborth H., Kemnitz A., Möller M. An upper bound for the minimum diameter of integral
point sets // Discrete \& Computational Geometry. — 1993. — т. 9, No 4. — с. 427—432.
\bibitem{piepmeyer1996maximum}
Piepmeyer L. The maximum number of odd integral distances between points in the plane //
Discrete \& Computational Geometry. — 1996. — т. 16, No 1. — с. 113—115.
\bibitem{kreisel2008heptagon}
Kreisel T., Kurz S. There are integral heptagons, no three points on a line, no four on a
circle // Discrete \& Computational Geometry. — 2008. — т. 39, No 4. — с. 786—790.
\bibitem{kurz2013constructing}
Constructing 7-clusters / S. Kurz [и др.] // Serdica Journal of Computing. — 2014. — т. 8,
No 1. — с. 47—70. — arXiv: 1312.2318.
\bibitem{solymosi2003note}
Solymosi J. Note on integral distances // Discrete \& Computational Geometry. — 2003. —
т. 30, No 2. — с. 337—342.
\bibitem{my-pps-linear-bound-2019}
Avdeev N. On existence of integral point sets and their diameter bounds. — 2019. — arXiv:
1906.11926.
\bibitem{bat2018number}
Bat-Ochir G. On the number of points with pairwise integral distances on a circle // Discrete
Applied Mathematics. — 2018. — т. 254. — с. 17—32.
\bibitem{brass2006research}
Brass P., Moser W. O., Pach J. Research problems in discrete geometry. — Springer Science
\& Business Media, 2006.
\bibitem{guy2013unsolved}
Guy R. Unsolved problems in number theory. т. 1. — Springer Science \& Business Media,
2013.
\bibitem{our-pmm-2018}
Авдеев Н. Н. Об отыскании целоудалённых множеств специального вида // Актуальные
проблемы прикладной математики, информатики и механики - сборник трудов Международной
научной конференции. — Научно-исследовательские публикации, 2018. — с. 492—498.
\bibitem{our-ped-2018}
Авдеев Н. Н. On integral point sets in special position // Некоторые вопросы анализа,
алгебры, геометрии и математического образования: материалы международной молодежной
научной школы «Актуальные направления математического анализа и смежные вопросы». —
2018. — т. 8. — с. 5—6.
\bibitem{kurz2005characteristic}
Kurz S. On the characteristic of integral point sets in $E^m$ // Australasian Journal of
Combinatorics. — 2006. — т. 36. — с. 241. — arXiv: math/0511704.
\bibitem{nozaki2013lower}
Nozaki H. Lower bounds for the minimum diameter of integral point sets // Australasian
Journal of Combinatorics. — 2013. — т. 56. — с. 139—143.
\bibitem{smurov1998stripcoverings}
Смуров М., Спивак А. Покрытия полосками // Квант. — 1998. — No 5. — с. 6.
\end{thebibliography}