### 2025-02-25 AvdMomZvol_rails/Article.tex
[rails] Edits from Zvolinsky: group the pics

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian,english]{babel}
\usepackage{amsmath,amsfonts,amssymb,euscript,graphicx,wrapfig,multirow}
\usepackage{dsfont}
\usepackage{amsthm}
\inputencoding{utf8}
%\bibliographystyle{unsrt}
\textheight=240mm \textwidth=170mm
\hoffset=-17mm
\voffset=-17mm

\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
\addbibresource{../common/notmy.bib}
\addbibresource{../common/my.bib}

\usepackage{../../../biblatex2bibitem/biblatex2bibitem}

%\usepackage{cite}

\usepackage{hyperref}

\makeatletter
%\renewcommand{\fnum@figure}{Figure \thefigure}
\renewcommand{\@biblabel}[1]{#1.}
\makeatother

\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}
\theoremstyle{dfn}
\newtheorem{dfn}{Definition}
\theoremstyle{remark}
\newtheorem{remark}{Remark}

\begin{document}
%\renewcommand\refname{\centering References}

% Переключаем язык на английский.
% Очень полезно как в плане типографики (в том числе расстановки переносов),
% так и в плане того, что не надо переименовывать "Рисунок" в "Figure"
\selectlanguage{english}

\title{
	On particular diameter bounds for integral point sets in higher dimensions
	\footnote{
		This work was carried out at Voronezh State University and supported by Russian Science
		Foundation grant 19-11-00197.
	}
}

%% Прекрасно понимаю, что следующая команда - дичь и вордописчество, но время не ждёт, время жмёт
\author{
	Avdeev N.N.
	\footnote{nickkolok@mail.ru, avdeev@math.vsu.ru}
	, Zvolinskii R.E., Momot E.A.
	\\
	\\
	\emph{Voronezh State University}
}

\maketitle

\paragraph{Absctract.}
	An integral point set $\mathcal{P}$ is a set of $n$ points in
	the $m$-dimensional Euclidean space $\mathbb{R}^{m}$ with pairwise
	integral distances,
	such that $\mathcal{P}$ is not contained in a $(m-1)$-dimensional hyperplane.
	In the present paper we discuss some classes of planar integral point sets ($m=2$);
	we mostly focus on the subsets of a union of two straight lines:
	facher, rails, scissors and pyramid sets.
	We use rails and pyramid sets to construct lower bounds for a diameter
	of integral point sets in higher dimensions.

\section{Introduction}
\label{sec:intro}

The study of integral point sets belongs to the classical analysis.
It is motivated by both pure theoretical aspects~\cite{ascher2020erdos,solymosi2010question}
and applications~\cite{kurz2007enumeration}.

\begin{dfn}\label{dfn1}
	A planar integral point set (PIPS) is a set $\mathcal{P}$
	of non-collinear points in the plane $\mathbb{R}^{2}$ such that
	for any pair of points $P_{1}, P_{2} \in \mathcal{P}$
	the Euclidean distance $|P_{1}P_{2}|$
	between points $P_{1}$ and $P_{2}$ is integral.
\end{dfn}

Definition \ref{dfn1} can be generalized as follows:

\begin{dfn}
	An integral point set $\mathcal{P}$ is a set of $n$ points in
	the $m$-dimensional Euclidean space $\mathbb{R}^{m}$ with pairwise
	integral distances,
	such that $\mathcal{P}$ is not contained in a $(m-1)$-dimensional hyperplane.
\end{dfn}

How can we characterize an integral point set?
First of all, we can look at its dimension;
then, we can find its cardinality, which is always finite~\cite{anning1945integral,erdos1945integral};
finally, we can compute the diameter of a finite point set,
which is naturally defined as follows:

\begin{dfn}
	The diameter of an integral point set $\mathcal{P}$ is defined as
	\begin{equation}
		\operatorname{diam(\mathcal{P})} = \underset{P_{1}, P_{2} \in
		\mathcal{P}}{\max} |P_{1}P_{2}|
		.
	\end{equation}
\end{dfn}

Every integral point set has also a characteristic~\cite{kemnitz1988punktmengen,kurz2005characteristic}.

\begin{dfn}
	The characteristic of a PIPS $\mathcal{P}$ is a squarefree number $q$
	such that the area of any triangle $M_1M_2M_3$, $\{M_1,M_2,M_3\}\subset \mathcal{P}$,
	is commensurable with $\sqrt{q}$.
\end{dfn}

The following notion was introduced in~\cite{kurz2008bounds}.

\begin{dfn}
	The function $d(m, n)$ is the minimal possible diameter of
	an integral point set $\mathcal{P}$ of $n$ points in
	$m$-dimensional Euclidean space $\mathbb{R}^{m}$.
\end{dfn}

The computation of exact values of $d(m,n)$ is a difficult problem.
In 2003, Solymosi proved~\cite{solymosi2003note} that $d(2,n) \geq cn$ for some constant $c$;
now it is known~\cite{my-pps-linear-bound-2019} that $c>5/11$.
The most recent advances for higher dimensions can be found in~\cite{nozaki2013lower}.
So, even finding the estimates is not easy.

A long list of known bounds for $d(m, n)$ and some exact values
can be found in~\cite[Theorem 1]{kurz2008bounds} or in~\cite{our-vmmsh-2018}.
Below we will discuss the following ones presented at~\cite{kurz2008bounds}:
\begin{align}
	d(m, 2m + 1) \leq 8  \label{eq:d_m_2m+1}\\
	d(m, 2m + 2) \leq 13 \label{eq:d_m_2m+2}\\
	d(m, 3m) \leq 109    \label{eq:d_m_3m}
\end{align}
and the following theorem \cite[Theorem 2.1]{kurz2008bounds}.

\begin{theorem}
	\label{thm:Kurz_blowup}
	Let $\mathcal{P}$ be a planar integral point set consisting of
	$n - 2$ points on the line $l_1$ and two points $P_{1}$ and $P_{2}$ on a
	parallel line $l_2$ with distance $r$ between $l_{1}$ and $l_{2}$. If there
	exist positive integers $v$, $w$ with $f^{2} + v^{2}
	= w^{2}$ and $v < 2r$, where $|P_{1}P_{2}| = f$,
	then

	\begin{equation}\label{formula1}
		d(m, n + 2(m - 2)) \leq \max(w, \operatorname{diam}(\mathcal{P}))
	\end{equation}

\end{theorem}

In this paper we further study the various features of IPS.
In Section~\ref{sec:classif}, we discuss the classification of planar integral points sets;
in Sections~\ref{sec:bounds_rails}~and~\ref{sec:bounds_pyramid}, we present some upper bounds for $d(m,n)$
based on planar integral point sets of special types
and provide some general constructions of such bounds.
All the bounds, like~\eqref{eq:d_m_2m+1}--\eqref{eq:d_m_3m}, are of the form $d(m, km+p)\leq c_{k,p}$.
%In particular, we improve the bound~\eqref{eq:d_m_3m} to
%\begin{equation}
%	d(m, 3m) \leq d(m, 3m + 1) \leq 56
%\end{equation}
%and present a new bound
%\begin{equation}
%	d(m, 4m + 1) \leq 1260
%	.
%\end{equation}

\section{Classification of planar integral point sets}
\label{sec:classif}

\subsection{Integral point sets situated on two straight lines}

\begin{dfn}
	A planar integral point sets of $n$ points with $n-1$ points on a straight line is called
	a \textit{facher} set.
\end{dfn}
Most of the examples of planar integral point sets turns to be facher sets.
In~\cite{antonov2008maximal}, facher sets with characteristic 1 are called \textit{semi-crabs}.
For $9 \leq n \leq 122$, the diameter $d(2,n)$ is attained on a facher point set~\cite{kurz2008minimum}.
Facher sets have relatively simple structure and thus
they are rather well-investigated~\cite{zvolinsky2021facher,lushina2021vestnikpmm}.

Non-facher integral point sets situated on two straight lines
can be further classificated in the following three cases:

\begin{dfn}
	A planar integral point set situated on two parallel straight lines
	is called a \textit{rails} set.
\end{dfn}

Among the rails sets, sets with 2 points on one line and all the other on another line prevail.
In Section~\ref{sec:bounds_rails} below, some examples of such sets are discussed;
they include an example of rails sets with 2 points on one line and 38 point on the other line
(Fig.~\ref{picture_40.png})
and
an example of rails set with 4 points on one line and 4 points on the other
(Fig.~\ref{picture.png}).
We still do not know whether there are rails set with 4 points on one line and 5 points on the other;
and the same for rails sets with 3 points on one line and 9 points on the other.

For the sake of brevity, we will hereafter use the notation
$\sqrt{p}/q * \{ (x_1,y_1), ...,$ $ (x_n, y_n)  \}$
from~\cite{our-ped-2018,our-pmm-2018,our-vmmsh-2018},
which means that each abscissa is multiplied by $1/q$
and each ordinate is multiplied by $\sqrt{p}/q$,  that is
$$
	\sqrt{p}/q * \{ (x_1,y_1), ..., (x_n, y_n)  \}
	=
	\left\{ \left(\frac{x_1}{q},\frac{y_1\sqrt{p}}{q}\right), ..., \left(\frac{x_n}{q},   \frac{y_n\sqrt{p}}{q}\right)  \right\}.
$$

On Fig.~\ref{11_2432_255255_a0a1b7f805cfc93080b4608505a0f45d_rails_3.png} we show the following rails set with 3 points on one line and 8 points on the other:
\begin{multline*}
\mathcal{P}_{3,8}=\sqrt{255255}/2*\{
	( 1767 ; -3);
	( 2791 ; -3);
	( 4071 ; -3);
\\
	( -306 ; 0);
	( 0 ; 0);
	( 1798 ; 0);
	( 2304 ; 0);
	( 2760 ; 0);
	( 3534 ; 0);
	( 4040 ; 0);
	( 4558 ; 0)
\}
\end{multline*}
It's notable that $\operatorname{char} \mathcal{P}_{3,8}=255255 = 3 \cdot 5 \cdot 7 \cdot 11 \cdot 13 \cdot 17$
is a product of the first 6 odd primes,
and the distances between the three points on the lower line are $512=2^{9}$, $640=2^7 \cdot 5$ and
$512+640 = 1152 = 2^7 \cdot 3^2$.

\begin{figure}[h!]
	\center{\includegraphics[width=0.9\linewidth]{11_2432_255255_a0a1b7f805cfc93080b4608505a0f45d_rails_3.png}}
	\parbox{0.9\linewidth}{\caption{PIPS of cardinality 11 and diameter 2423}
	\label{11_2432_255255_a0a1b7f805cfc93080b4608505a0f45d_rails_3.png}}
\end{figure}

Rails sets with 4 and 4 points on the lines are somewhat rather common;
an example of such a set is shown on Fig.~\ref{8_1111_770_5666c83a03a80e1d3f8f8e8186a145f6__rails_4.png}:
\begin{equation*}
	\sqrt{770}/1*\{
		( 0 ; 0);
		( 380 ; 0);
		( 731 ; 0);
		( 1111 ; 0);
		( 354 ; -12);
		( 377 ; -12);
		( 734 ; -12);
		( 757 ; -12)%;
	\}
\end{equation*}

\begin{figure}[h!]
	\center{\includegraphics[width=0.9\linewidth]{8_1111_770_5666c83a03a80e1d3f8f8e8186a145f6__rails_4.png}}
	\parbox{0.9\linewidth}{\caption{PIPS of cardinality 8 and diameter 1111}
	\label{8_1111_770_5666c83a03a80e1d3f8f8e8186a145f6__rails_4.png}}
\end{figure}

\begin{dfn}
	A planar integral point set situated on two perpendicular straight lines
	is called a \textit{cross} set.
\end{dfn}
Every cross set has characteristic 1;
in~\cite{antonov2008maximal}, cross sets with only 2 points out of one of the lines that have two axes of symmetry are called \textit{crabs}.

\begin{dfn}
	A planar integral point set situated on two straight lines
	that are not parallel nor perpendicular,
	is called a \textit{scissors} set.
\end{dfn}

There is an important subclass of scissors sets.

\begin{dfn}
	A scissors set with an axis of symmetry,
	which is the angle bisector for the straight lines,
	is called a \textit{pyramid} set.
\end{dfn}

Examples of pyramid sets are given in Section~\ref{sec:bounds_pyramid} on
Fig.~\ref{picture_56.png} and Fig.~\ref{picture_1260.png}.
We don't have any examples of pyramid sets with cardinality more than 9.

It still remains unknown whether a scissors set can have another axis of symmetry;
on Fig.~\ref{15_3364_91_30e991f82a54fb9caeca5b6a6776b981_X.png} the following scissors set with central symmetry is shown:
\begin{multline*}
	\sqrt{91}/1*\{
	( 0 ; 0);
	( 792 ; 120);
	( -792 ; -120);
\\
	( \pm256 ; 0);
	( \pm646 ; 0);
	( \pm693 ; 0);
	( \pm1152 ; 0);
	( \pm1242 ; 0);
	( \pm1682 ; 0)
\}
\end{multline*}

\subsection{Other integral point sets}

\begin{dfn}
	A planar integral point set that is situated on a circle is called a \textit{circular}
	point set.
\end{dfn}

Circular sets are very important examples
of integral point sets~\cite{harborth1993upper,piepmeyer1996maximum,bat2018number};
for example, the best known upper bound for $d(2,n)$
is attained on a circular set.
Moreover, the first occurrence of integral point sets in modern literature (or at least the first one that the authors are aware of) is Anning's article from 1915 which focuses exactly on circular integral point sets~\cite{anning1915discussions}.
For the most recent advances in circular IPS, we refer the reader to~\cite{undrakh2022refinement}.

\begin{dfn}
	A planar integral point set that is situated on the conjunction of a circle with its center,
	is called a \textit{centered-circular} point set.
\end{dfn}

\begin{remark}
	Every centered-circular point set has characteristic 1.
	Indeed, let $\mathcal{P} = \{M_0, M_1,$ $ M_2, M_3, ..., M_k\}$ be a centered-circular point set
	with $M_0$ being the center.
	By the Law of sines, in the triangle $M_1 M_2 M_3$
	\begin{equation}
		\sin \angle M_1 M_2 M_3 = \frac{|M_1M_3|}{2R}
		,
	\end{equation}
	where $R = |M_0 M_1| $ is the radius of the triangle's circumcircle.
	Then the area of the triangle $M_1 M_2 M_3$ is
	\begin{equation}
		S = \frac12 |M_2 M_1| \cdot |M_2 M_3| \cdot \sin \angle M_1 M_2 M_3 =
		\frac12 |M_2 M_1| \cdot |M_2 M_3| \cdot \frac{|M_1M_3|}{2|M_0 M_1|} \in\mathbb Q
		.
	\end{equation}
\end{remark}

\begin{figure}[htbp]
	\includegraphics[width=.42\linewidth]{17_2210_1_acc4a56887abb3c3a315712ea083f0e7_centered_circle.png}
	\hfill
	\includegraphics[width=.545\linewidth]{15_3364_91_30e991f82a54fb9caeca5b6a6776b981_X.png}
	\\
	\parbox{.37\linewidth}{\caption{A centered-circular PIPS of cardinality 17 and diameter 2210}
	\label{17_2210_1_acc4a56887abb3c3a315712ea083f0e7_centered_circle.png}}
	\hfill
	\parbox{.5\linewidth}{\caption{A scissors PIPS of cardinality 15 and diameter 3364}
	\label{15_3364_91_30e991f82a54fb9caeca5b6a6776b981_X.png}}
\end{figure}

On Fig.~\ref{17_2210_1_acc4a56887abb3c3a315712ea083f0e7_centered_circle.png} the following centered-circular PIPS is shown:
\begin{multline*}
	\sqrt{1}/1105*\{
	( 0 ; 0);
	( \pm1221025 ; 0);
	( 113953 ; 1215696);
	( -113953 ; -1215696);
	\\
	( 341887 ; 1172184);
	( -341887 ; -1172184);
	( 680225 ; 1014000);
	( -680225 ; -1014000);
	\\
	( 782977 ; -936936);
	( -782977 ; 936936);
	( 859775 ; -867000);
	( -859775 ; 867000);
	\\
	( 1073057 ; 582624);
	( -1073057 ; -582624);
	( 1198975 ; 231000);
	( -1198975 ; -231000);
\}
\end{multline*}

Most of the known planar integral point sets fall into these six classes;
however, there are some sophisticated examples which do not~\cite{kreisel2008heptagon,kurz2013constructing,avdeev2021particular}.
A lower bound for the diameter of a PIPS with no collinear triples (so-called PIPS \textit{in semi-general position})
is presented in~\cite{my-semi-general-5-4-bound-2019}.
It is important to note that the cardinality of an IPS in semi-general position is bounded
not only by its diameter (i.e. the maximal distance between its points),
but also by the minimal distance.
Solymosi established such a bound~\cite{solymosi2003note};
in particular, he proved that if an IPS in semi-general position has distance 2,
its cardinality cannot exceed 7.
Recently, this result has been refined to 5 points~\cite{lushina2024cardinality}.

%TODO: pictures!

\section{Bounds based on rails sets}
\label{sec:bounds_rails}

Let $i = \overline{1, k}$ denote the enumeration of all $i$
from $1$ to $k$.
Theorem \cite[Theorem 2.1]{kurz2008bounds} can be generalized as follows:

\begin{theorem}
	\label{thm:rails_blowup}
	Let $\mathcal{P}$ be a planar integral point set consisting of
	$n - k$ points on the line $l_1$ and $k$ points $P_1$, $P_2$, ..., $P_k$ on a
	parallel line $l_2$ with distance $r$ between $l_1$ and $l_2$. If there
	exist positive integers $v$, $w_{ij}$, with $f_{ij}^{2} + v^{2}
	= w_{ij}^{2}$ and $v < 2r$, where $i = \overline{1, k - 1}$, $j =
	\overline{i + 1, k}$, $|P_{i}P_{j}| = f_{ij}$,
	then

	\begin{equation}
		d(m, n + k(m - 2)) \leq \max(w_{1k}, \operatorname{diam(\mathcal{P})})
	\end{equation}

\end{theorem}

Below we give the examples of planar integral point sets and the corresponding
estimates of the function $d(m, n)$ for $n = 2m + k$, $3 \leq k \leq 36$.

Below $f$ and $w$ are those from Theorem~\ref{thm:Kurz_blowup}.

\begin{itemize}
\setlength{\itemsep}{-1mm}

\item
$\mathcal{P}_7 = \sqrt{315}/{2} * \{ (13, 1),
(29, 1),
(0, 0),
(10, 0),
(16, 0),
(26, 0),
(32, 0)\}
$
,

$f = 8$, $v = 6$, $w = 10$, $\operatorname{diam(\mathcal{P}_7)} = 17$,
which gives $d(m, 2m + 3) \leq 17$.

\item
$\mathcal{P}_8 = \mathcal{P}_7 \cup \sqrt{315}/2 * \{(42, 0)\}
$
(Figure~\ref{picture_d21.png}),

$f = 8$, $v = 6$, $w = 10$, $\operatorname{diam(\mathcal{P}_8)} = 21$,
which gives $d(m, 2m + 4) \leq 21$.

\item
$\mathcal{P}_9 = \sqrt{70}/{1} * \{ (\pm 44, 12),
(-78 , 0),
(\pm 62, 0),
(\pm 55 , 0),
(\pm 10 , 0)\}
$
,

$f = 88$, $v = 66$, $w = 110$, $\operatorname{diam(\mathcal{P}_9)} = 158$,

which gives $d(m, 2m + 5) \leq 158$.

\item
$\mathcal{P}_{10} = \mathcal{P}_9 \cup \sqrt{70}/{1} * \{(78, 0)\}
$
(Figure~\ref{picture_3.png}),

$f = 88$, $v = 66$, $w = 110$, $\operatorname{diam(\mathcal{P}_{10})} = 158$,

which gives $d(m, 2m + 6) \leq 158$.

\begin{figure}[h!]
	\includegraphics[width=.48\linewidth]{picture_d21.png}
	\hfill
	\includegraphics[width=.48\linewidth]{picture_3.png}
	\\
	\parbox{.48\linewidth}{\caption{PIPS $\mathcal{P}_8$ of cardinality 8 and diameter 21}
		\label{picture_d21.png}}
	\hfill
	\parbox{.48\linewidth}{\caption{PIPS $\mathcal{P}_{10}$ of cardinality 10 and diameter 158}
		\label{picture_3.png}}
\end{figure}

\item
$\mathcal{P}_{11} = \sqrt{70}/{2} * \{ (\pm 99, 24),
(\pm 207 , 0),
(\pm 145 , 0),
(\pm 63 , 0),
(\pm 25 , 0),
(-297 , 0)\}
$
,

$f = 99$, $v = 20$, $w = 101$, $\operatorname{diam(\mathcal{P}_{11})} = 252$,

which gives $d(m, 2m + 7) \leq 252$.

\item
$\mathcal{P}_{12} = \mathcal{P}_{11} \cup \sqrt{70}/{2} * \{(297, 0)\}
$
(Figure~\ref{picture_5.png}),

$f = 99$, $v = 20$, $w = 101$, $\operatorname{diam(\mathcal{P}_{12})} = 297$,

which gives $d(m, 2m + 8) \leq 297$.

\begin{figure}[h!]
	\includegraphics[width=.48\linewidth]{picture_5.png}
	\hfill
	\includegraphics[width=.48\linewidth]{picture_7.png}
	\\
	\parbox{.48\linewidth}{\caption{PIPS $\mathcal{P}_{12}$ of cardinality 12 and diameter 297}
		\label{picture_5.png}}
	\hfill
	\parbox{.48\linewidth}{\caption{PIPS $\mathcal{P}_{14}$ of cardinality 14 and diameter 1273}
		\label{picture_7.png}}
\end{figure}

\item
$\mathcal{P}_{13} = \sqrt{19019}/{2} * \{ (\pm 200, 3),
(\pm 873 , 0),
(\pm 615 , 0),
(\pm 377 , 0),
(\pm 215 , 0),
(\pm 23 , 0),
$

$
(-1273 , 0)\}
$
,

$f = 200$, $v = 45$, $w = 205$, $\operatorname{diam(\mathcal{P}_{13})} = 1073$,

which gives $d(m, 2m + 9) \leq 1073$.

\item
$\mathcal{P}_{14} = \mathcal{P}_{13} \cup \sqrt{19019}/{2} * \{(1273, 0)\}
$
(Figure~\ref{picture_7.png}),

$f = 200$, $v = 45$, $w = 205$, $\operatorname{diam(\mathcal{P}_{14})} = 1273$,

which gives $d(m, 2m + 10) \leq 1273$.

\item
$\mathcal{P}_{15} = \sqrt{385}/{2} * \{ (\pm 1105, 48),
(\pm 1587 , 0),
(\pm 1269 , 0),
(\pm 763 , 0),
(\pm 623 , 0),
(\pm 529 , 0),
$

$
(\pm 339 , 0),
(-2189 , 0)\}
$
,

$f = 1105$, $v = 300$, $w = 1145$, $\operatorname{diam(\mathcal{P}_{15})} = 1888$,

which gives $d(m, 2m + 11) \leq 1888$.

\item
$\mathcal{P}_{16} = \mathcal{P}_{15} \cup \sqrt{385}/{2} * \{(2189, 0)\}
$
(Figure~\ref{picture_9.png}),

$f = 1105$, $v = 300$, $w = 1145$, $\operatorname{diam(\mathcal{P}_{16})} = 2189$,

which gives $d(m, 2m + 12) \leq 2189$.

\begin{figure}[h!]
\center{\includegraphics[width=0.75\linewidth]{picture_9.png}}
\parbox{0.75\linewidth}{\caption{PIPS $\mathcal{P}_{16}$ of cardinality 16 and diameter 2189}
\label{picture_9.png}}
\end{figure}

\item
$\mathcal{P}_{17} = \sqrt{154}/{1} * \{ (\pm 874, 60),
(\pm 1376 , 0),
(\pm 1036 , 0),
(\pm 899 , 0),
(\pm 849 , 0),
(\pm 613 , 0),
$

$
(\pm 576 , 0),
(\pm 100 , 0),
(-1848 , 0)\}
$
,

$f = 1748$, $v = 336$, $w = 1780$, $\operatorname{diam(\mathcal{P}_{17})} = 3224$,

which gives $d(m, 2m + 13) \leq 3224$.

\item
$\mathcal{P}_{18} = \mathcal{P}_{17} \cup \sqrt{154}/{1} * \{(1848, 0)\}
$
(Figure~\ref{picture_14.png}),

$f = 1748$, $v = 336$, $w = 1780$, $\operatorname{diam(\mathcal{P}_{18})} = 3696$,

which gives $d(m, 2m + 14) \leq 3696$.

\begin{figure}[h!]
\center{\includegraphics[width=0.8\linewidth]{picture_14.png}}
\parbox{0.85\linewidth}{\caption{PIPS $\mathcal{P}_{18}$ of cardinality 18 and diameter 3696}
\label{picture_14.png}}
\end{figure}

\item
$\mathcal{P}_{19} = \mathcal{P}_{18} \cup \sqrt{154}/{1} * \{(-3293, 0)\}
$
,

$f = 1748$, $v = 336$, $w = 1780$, $\operatorname{diam(\mathcal{P}_{19})} = 5141$,

which gives $d(m, 2m + 15) \leq 5141$.

\item
$\mathcal{P}_{20} = \mathcal{P}_{19} \cup \sqrt{154}/{1} * \{(3293, 0)\}
$
(Figure~\ref{picture_16.png}),

$f = 1748$, $v = 336$, $w = 1780$, $\operatorname{diam(\mathcal{P}_{20})} = 6586$,

which gives $d(m, 2m + 16) \leq 6586$.

\begin{figure}[h!]
\center{\includegraphics[width=1\linewidth]{picture_16.png}}
\parbox{1\linewidth}{\caption{PIPS $\mathcal{P}_{20}$ of cardinality 20 and diameter 6586}
\label{picture_16.png}}
\end{figure}

\item
$\mathcal{P}_{21} = \sqrt{154}/{1} * \{ (\pm 2622, 180),
(\pm 5544 , 0),
(\pm 4128 , 0),
(\pm 3108 , 0),
(\pm 2697 , 0),
$

$
(\pm 2547 , 0),
(\pm 1839 , 0),
(\pm 1728 , 0),
(\pm 904 , 0),
(\pm 300 , 0),
(-6148 , 0)\}
$
,

$f = 5244$, $v = 1008$, $w = 5340$, $\operatorname{diam(\mathcal{P}_{21})} = 11692$,

which gives $d(m, 2m + 17) \leq 11692$.

\item
$\mathcal{P}_{22} = \mathcal{P}_{21} \cup \sqrt{154}/{1} * \{(6148, 0)\}
$
(Figure~\ref{picture_18.png}),

$f = 5244$, $v = 1008$, $w = 5340$, $\operatorname{diam(\mathcal{P}_{22})} = 12296$,

which gives $d(m, 2m + 18) \leq 12296$.

\item
$\mathcal{P}_{23} = \mathcal{P}_{22} \cup \sqrt{154}/{1} * \{(-9879 , 0)\}
$
,

$f = 5244$, $v = 1008$, $w = 5340$, $\operatorname{diam(\mathcal{P}_{23})} = 16027$,

which gives $d(m, 2m + 19) \leq 16027$.

\begin{figure}[h!]
\center{\includegraphics[width=0.75\linewidth]{picture_18.png}}
\parbox{0.75\linewidth}{\caption{PIPS $\mathcal{P}_{22}$ of cardinality 22 and diameter 12296}
\label{picture_18.png}}
\end{figure}

\item
$\mathcal{P}_{24} = \mathcal{P}_{23} \cup \sqrt{154}/{1} * \{ (9879, 0)\}
$
(Figure~\ref{picture_20.png}),

$f = 5244$, $v = 1008$, $w = 5340$, $\operatorname{diam(\mathcal{P}_{24})} = 19758$,

which gives $d(m, 2m + 20) \leq 19758$.

\item
$\mathcal{P}_{25} = \sqrt{154}/{1} * \{ (\pm 5244, 360),
(\pm 12296 , 0),
(\pm 11088 , 0),
(\pm 8579 , 0),
(\pm 8256 , 0),
$

$
(\pm 6216 , 0),
(\pm 5394 , 0),
(\pm 5094 , 0),
(\pm 3678 , 0),
(\pm 3456 , 0),
(\pm 1808 , 0),
(\pm 600 , 0),
$

$
(-19758 , 0)\}
$
,

$f = 10488$, $v = 1015$, $w = 10537$, $\operatorname{diam(\mathcal{P}_{25})} = 32054$,

which gives $d(m, 2m + 21) \leq 32054$.

\begin{figure}[h!]
\center{\includegraphics[width=0.85\linewidth]{picture_20.png}}
\parbox{0.85\linewidth}{\caption{PIPS $\mathcal{P}_{24}$ of cardinality 24 and diameter 19758}
\label{picture_20.png}}
\end{figure}

\item
$\mathcal{P}_{26} = \mathcal{P}_{25} \cup \sqrt{154}/{1} * \{(19758 , 0)\}
$
(Figure~\ref{picture_22.png}),

$f = 10488$, $v = 1015$, $w = 10537$, $\operatorname{diam(\mathcal{P}_{26})} = 39516$,

which gives $d(m, 2m + 22) \leq 39516$.

\item
$\mathcal{P}_{27} = \sqrt{154}/{1} * \{ (\pm 36708, 2520),
(\pm 116058 , 0),
(\pm 86072 , 0),
(\pm 77616 , 0),
(\pm 60053 , 0),
$

$
(\pm 57792 , 0),
(\pm 43512 , 0),
(\pm 37758 , 0),
(\pm 35658 , 0),
(\pm 25746 , 0),
(\pm 24192 , 0),
$

$
(\pm 12656 , 0),
(\pm 4200 , 0),
(-138306 , 0)\}
$
,

$f = 73416$, $v = 2710$, $w = 73466$, $\operatorname{diam(\mathcal{P}_{27})} = 254364$,

which gives $d(m, 2m + 23) \leq 254364$.

\begin{figure}[h!]
\center{\includegraphics[width=1\linewidth]{picture_22.png}}
\parbox{1\linewidth}{\caption{PIPS $\mathcal{P}_{26}$ of cardinality 26 and diameter 39516}
\label{picture_22.png}}
\end{figure}

\item
$\mathcal{P}_{28} = \mathcal{P}_{27} \cup \sqrt{154}/{1} * \{(138306, 0)\}
$
(Figure~\ref{picture_24.png}),

$f = 73416$, $v = 2710$, $w = 73466$, $\operatorname{diam(\mathcal{P}_{28})} = 276612$,

which gives $d(m, 2m + 24) \leq 276612$.

\item
$\mathcal{P}_{29} = \sqrt{154}/{1} * \{ (\pm 36708, 2520),
(\pm 138306 , 0),
(\pm 116058 , 0),
(\pm 86072 , 0),
(\pm 77616 , 0),
$

$
(\pm 60053 , 0),
(\pm 57792 , 0),
(\pm 43512 , 0),
(\pm 37758 , 0),
(\pm 35658 , 0),
(\pm 25746 , 0),
$

$
(\pm 24192 , 0),
(\pm 12656 , 0),
(\pm 4200 , 0),
(-199863 , 0)\}
$
,

$f = 73416$, $v = 2710$, $w = 73466$, $\operatorname{diam(\mathcal{P}_{29})} = 338169$,

which gives $d(m, 2m + 25) \leq 338169$.

\begin{figure}[h!]
\center{\includegraphics[width=1\linewidth]{picture_24.png}}
\parbox{1\linewidth}{\caption{PIPS $\mathcal{P}_{28}$ of cardinality 28 and diameter 276612}
\label{picture_24.png}}
\end{figure}

\item
$\mathcal{P}_{30} = \mathcal{P}_{29} \cup \sqrt{154}/{1} * \{(199863 , 0)\}
$
(Figure~\ref{picture_26.png}),

$f = 73416$, $v = 2710$, $w = 73466$, $\operatorname{diam(\mathcal{P}_{30})} = 399726$,

which gives $d(m, 2m + 26) \leq 399726$.

\begin{figure}[h!]
\center{\includegraphics[width=1\linewidth]{picture_26.png}}
\parbox{1\linewidth}{\caption{PIPS $\mathcal{P}_{30}$ of cardinality 30 and diameter 399726}
\label{picture_26.png}}
\end{figure}

\item
$\mathcal{P}_{31} = \sqrt{154}/{1} * \{ (\pm 580290, 0),
(\pm 430360, 0),
(\pm 388080, 0),
(\pm 300265, 0),
$

$
(\pm 288960, 0),
(\pm 217560, 0),
(\pm 188790, 0),
(\pm 183540, 12600),
(\pm 178290, 0),
$

$
(\pm 128730, 0),
(\pm 120960, 0),
(\pm 63280, ),
(\pm 21000, 0),
(\pm 11224, 0),
$

$
(-1033912, 0),
(-999315, 0),
(-691530  , 0)\}
$
,

$f = 367080$, $v = 3206$, $w = 367094$, $\operatorname{diam(\mathcal{P}_{31})} = 1614202$,

which gives $d(m, 2m + 27) \leq 1614202$.

\item
$\mathcal{P}_{32} = \mathcal{P}_{31} \cup \sqrt{154}/{1} * \{(691530  , 0)\}
$
(Figure~\ref{picture_32.png}),

$f = 367080$, $v = 3206$, $w = 367094$, $\operatorname{diam(\mathcal{P}_{32})} = 1725442$,

which gives $d(m, 2m + 28) \leq 1725442$.

\begin{figure}[h!]
\center{\includegraphics[width=1\linewidth]{picture_32.png}}
\parbox{1\linewidth}{\caption{PIPS $\mathcal{P}_{32}$ of cardinality 32 and diameter 1725442}
\label{picture_32.png}}
\end{figure}

\item
$\mathcal{P}_{33} = \mathcal{P}_{32} \cup \sqrt{154}/{1} * \{(999315, 0)\}
$
,

$f = 367080$, $v = 3206$, $w = 367094$, $\operatorname{diam(\mathcal{P}_{33})} = 2033227$,

which gives $d(m, 2m + 29) \leq 2033227$.

\item
$\mathcal{P}_{34} = \mathcal{P}_{33} \cup \sqrt{154}/{1} * \{(1033912, 0)\}
$
(Figure~\ref{picture_34.png}),

$f = 367080$, $v = 3206$, $w = 367094$, $\operatorname{diam(\mathcal{P}_{34})} = 2067824$,

which gives $d(m, 2m + 30) \leq 2067824$.

\begin{figure}[h!]
\center{\includegraphics[width=1\linewidth]{picture_34.png}}
\parbox{1\linewidth}{\caption{PIPS $\mathcal{P}_{34}$ of cardinality 34 and diameter 2067824}
\label{picture_34.png}}
\end{figure}

\item
$\mathcal{P}_{35} = \sqrt{154}/{1} * \{ (\pm 3997260, 0),
(\pm 2766120, 0),
(\pm 2321160, 0),
(\pm 1721440, 0),
$

$
(\pm 1552320, 0),
(\pm 1201060, 0),
(\pm 1155840, 0),
(\pm 870240, 0),
(\pm 755160, 0),
$

$
(\pm 734160, 50400),
(\pm 713160, 0),
(\pm 514920, 0),
(\pm 483840, 0),
(\pm 308175, 0),
$

$
(\pm 253120, 0),
(\pm 84000, 0),
(\pm 44896, 0),
(-4135648 , 0)\}
$
,

$f = 1468320$, $v = 12824$, $w = 1468376$, $\operatorname{diam(\mathcal{P}_{35})} = 8132908$,

which gives $d(m, 2m + 31) \leq 8132908$.

\item
$\mathcal{P}_{36} = \mathcal{P}_{35} \cup \sqrt{154}/{1} * \{(4135648, 0)\}
$
(Figure~\ref{picture_36.png}),

$f = 1468320$, $v = 12824$, $w = 1468376$, $\operatorname{diam(\mathcal{P}_{36})} = 8271296$,

which gives $d(m, 2m + 32) \leq 8271296$.

\begin{figure}[h!]
\center{\includegraphics[width=1\linewidth]{picture_36.png}}
\parbox{1\linewidth}{\caption{PIPS $\mathcal{P}_{36}$ of cardinality 36 and diameter 8271296}
\label{picture_36.png}}
\end{figure}

\item
$\mathcal{P}_{37} = \sqrt{154}/{1} * \{ (\pm 12991095, 0),
(\pm 8989890, 0),
(\pm 7543770, 0),
(\pm 5594680, 0),
$

$
(\pm 5045040, 0),
(\pm 3903445, 0),
(\pm 3756480, 0),
(\pm 3694845, 0),
(\pm 2828280, 0),
$

$
(\pm 2454270, 0),
(\pm 2386020, 163800),
(\pm 2317770, 0),
(\pm 1673490, 0),
(\pm 1572480, 0),
$

$
(\pm 822640, 0),
(\pm 484680, 0),
(\pm 273000, 0),
(\pm 145912, 0),
(-13440856, 0)\}
$
,

$f = 4772040$, $v = 41678$, $w = 4772222$, $\operatorname{diam(\mathcal{P}_{37})} = 26431951$,

which gives $d(m, 2m + 33) \leq 26431951$.

\item
$\mathcal{P}_{38} = \mathcal{P}_{37} \cup \sqrt{154}/{1} * \{(13440856, 0)\}
$
(Figure~\ref{picture_38.png}),

$f = 4772040$, $v = 41678$, $w = 4772222$, $\operatorname{diam(\mathcal{P}_{38})} = 26881712$,

which gives $d(m, 2m + 34) \leq 26881712$.

\begin{figure}[h!]
\center{\includegraphics[width=1\linewidth]{picture_38.png}}
\parbox{1\linewidth}{\caption{PIPS $\mathcal{P}_{38}$ of cardinality 38 and diameter 26881712}
\label{picture_38.png}}
\end{figure}

\item
$\mathcal{P}_{39} = \sqrt{154}/{1} * \{ (\pm 51964380, 0),
(\pm 35959560, 0),
(\pm 30175080, 0),
(\pm 22378720, 0),
$

$
(\pm 20180160, 0),
(\pm 15613780, 0),
(\pm 15025920, 0),
(\pm 14779380, 0),
(\pm 11313120, 0),
$

$
(\pm 9817080, 0),
(\pm 9544080, 655200),
(\pm 9271080, 0),
(\pm 6693960, 0),
(\pm 6289920, 0),
$

$
(\pm 4006275, 0),
(\pm 3290560, 0),
(\pm 1938720, 0),
(\pm 1092000, 0),
$

$
(\pm 583648, 0),
(-53763424 , 0)\}
$
,

$f = 19088160$, $v = 8738$, $w = 19088162$, $\operatorname{diam(\mathcal{P}_{39})} = 105727804$,

which gives $d(m, 2m + 35) \leq 105727804$.

\begin{figure}[h!]
	\begin{center}
	\includegraphics[width=1\linewidth]{picture_40.png}
	\parbox{1\linewidth}{\caption{PIPS of cardinality 40 and diameter 107526848}
	\label{picture_40.png}}
	\end{center}
\end{figure}

\item
$\mathcal{P}_{40} = \mathcal{P}_{39} \cup \sqrt{154}/{1} * \{(53763424 , 0)\}
$
(Figure~\ref{picture_40.png}),

$f = 19088160$, $v = 8738$, $w = 19088162$, $\operatorname{diam}(\mathcal{P}_{40}) = 107526848$,

which gives $d(m, 2m + 36) \leq 107526848$.

\end{itemize}

\begin{remark}
In all the examples above, the maximum in the right-hand side of (\ref{formula1})
is attained on the diameter of the system.
The following example shows that the maximum can be attained on $w$:
\end{remark}

\begin{figure}[h!]
	\includegraphics[width=.48\linewidth]{picture_10.png}
	\hfill
	\includegraphics[width=.48\linewidth]{picture.png}
	\\
	\parbox{.48\linewidth}{\caption{PIPS of cardinality 8 and diameter 162}
	\label{picture_10.png}}
	\hfill
	\parbox{.48\linewidth}{\caption{PIPS of cardinality 8 and diameter 409}
	\label{picture.png}}
\end{figure}

\begin{itemize}
\setlength{\itemsep}{-1mm}

\item
$\mathcal{P}=\sqrt{2}/{1} * \{ (\pm 76, 60),
(\pm 62 , 0),
(\pm 29 , 0),
(\pm 6 , 0)\}
$
(Figure~\ref{picture_10.png}),

$f = 152$, $v = 114$, $w = 190$, $\operatorname{diam(\mathcal{P})} = 162$.

We can obtain the estimate $d(m, 2m + 2) \leq 190$ (which
is weaker than the one given in~\eqref{eq:d_m_2m+2}).

\end{itemize}

Guy gives \cite[D 20]{guy2013unsolved} a PIPS of $8$ points located
on two parallel lines. The points of the set form a rectangle
(Figure~\ref{rectangle.pdf}):

\begin{figure}[htbp]
	\includegraphics[width=.6\linewidth]{rectangle.pdf}
	\hfill
	\includegraphics[width=.5\linewidth]{right_angle.pdf}
	\\
	\parbox{.65\linewidth}{\caption{Distances}
	\label{rectangle.pdf}}
	\hfill
	\parbox{.5\linewidth}{\caption{Contradiction}
	\label{right_angle.pdf}}
\end{figure}

\begin{itemize}
\setlength{\itemsep}{-1mm}

\item
$\mathcal{P}=\sqrt{1}/{1} * \{ (0, \pm 60),
(182 , \pm 60),
(209 , \pm 60),
(391 , \pm 60)\}
$
(Figure~\ref{picture.png}),

where $n = 8$, $\operatorname{diam(\mathcal{P})} = 409$. Using the ``blowing up''
construction, we obtain
\begin{equation}\label{result1}
d(m, 4m) \leq 409
\end{equation}
Figure~\ref{picture_11.pdf} shows an example for $m = 3$.

\begin{figure}[h!]
\center{\includegraphics[width=1\linewidth]{picture_11.pdf}}
\parbox{1\linewidth}{\caption{IPS of cardinality 12 and diameter 409}
\label{picture_11.pdf}}
\end{figure}

\end{itemize}

\section{Bounds based on pyramid sets}
\label{sec:bounds_pyramid}

\begin{theorem}
Let $\mathcal{P}$ be a planar integral point set consisting of $k$
points on the line $l_{1}$ and $k$ points on the line $l_{2}$. Besides, these
points are symmetric with respect to one of the bisectors of the angles
formed by the intersection of lines $l_{1}$ and $l_{2}$, then

\begin{equation}\label{formula2}
d(m, (k - \alpha)m + \alpha) \leq \operatorname{diam}(\mathcal{P}),
\end{equation}

where
\begin{equation*}
\alpha =
\begin{cases}
1, \text{when the intersection point} \in \mathcal{P}, \\
0, \text{when the intersection point} \notin \mathcal{P}.
\end{cases}
\end{equation*}

\end{theorem}

\begin{remark}
The angles can be acute or obtuse. Figure~\ref{right_angle.pdf} shows that the intersection angle of the lines $l_{1}$ and $l_{2}$ cannot be equal to ${\pi}/2$.
\end{remark}

Below we give the examples of planar integral point sets and the corresponding
estimates of the function $d(m, n)$ for $n = 3m + 1$ and $n = 4m + 1$.

\begin{figure}[htbp]
	\includegraphics[width=.48\linewidth]{picture_56.png}
	\hfill
	\includegraphics[width=.48\linewidth]{picture_1260.png}
	\\
	\parbox{.48\linewidth}{\caption{PIPS of cardinality 7 and diameter 56}
	\label{picture_56.png}}
	\hfill
	\parbox{.48\linewidth}{\caption{PIPS of cardinality 9 and diameter 1260}
	\label{picture_1260.png}}
\end{figure}

\begin{figure}[h!]
\center{\includegraphics[width=1\linewidth]{picture_12.pdf}}
\parbox{1\linewidth}{\caption{IPS of cardinality 10 and diameter 56}
\label{picture_12.pdf}}
\end{figure}

\begin{figure}[h!]
\center{\includegraphics[width=1\linewidth]{picture_1260_R3.pdf}}
\parbox{1\linewidth}{\caption{IPS of cardinality 13 and diameter 1260}
\label{picture_1260_R3.pdf}}
\end{figure}

\begin{itemize}
\setlength{\itemsep}{-1mm}

\item
$\mathcal{P}=\sqrt{3}/{2} * \{ (\pm 56, 0),
(14, 0),
(-34, 0),
(-10, 24),
(-21 , 35),
(35, -21)\}
$
(Figure~\ref{picture_56.png}),

where $n = 7$, $\operatorname{diam(\mathcal{P})} = 56$. Using the ``blowing up''
construction, we obtain
\begin{equation}\label{result2}
d(m, 3m + 1) \leq 56
\end{equation}
The resulting estimate improves the estimate for $d(m, 3m)$, which is presented
in \cite{kemnitz1988punktmengen}. Figure~\ref{picture_12.pdf} shows an example
for $m = 3$.

\item
$\mathcal{P}=\sqrt{39}/{8} * \{ (\pm 5040, 0),
(1911, 315),
(2352, 0),
(944, 0),
(336, 0),
(2940, -420),
(2044, 220),
$

$
(735, 1155)\}
$
(Figure~\ref{picture_1260.png}),

where $n = 9$, $\operatorname{diam(\mathcal{P})} = 1260$. Using the ``blowing up''
construction, we obtain
\begin{equation}\label{result3}
d(m, 4m + 1) \leq 1260
\end{equation}
Figure~\ref{picture_1260_R3.pdf} shows an example for $m = 3$.

However, we have to admit that both bounds which employ pyramid sets
are indeed weaker than the one
\begin{equation}
	d(m, m^2+m)\leq 17
\end{equation}
proved in~\cite{kurz2008bounds},
so the constructions of bounds based on pyramid sets tend to be only for the conceptual purpose.

\end{itemize}

\section{Final remarks}
All the discussed PIPSs were obtained through a combination of computer search and intuition of the authors;
so, the further search may lead to better bounds employing the same constructions.

There is still no general construction for a rails or scissors PIPS of arbitrary cardinality.
For rails PIPSs, we can conjecture that there exists a set of arbitrary cardinality, with 2 points on one line
and all the rest on the other;
for an example of cardinality 106, we refer the reader to~\cite{momot2022example}.
On the other hand, it is still unknown whether there are any rails PIPSs with 4 and 5 points on the lines.

As for today, we have found pyramid PIPS of cardinality at most 9.

The source code can be obtained at https://gitlab.com/Nickkolok/ips-algo .
The preprint of this article is arXiv:1909.10386 [math.CO]

The authors thank A.S. Usachev for his help in polishing the text of the paper.

Авдеев Николай Николаевич

Воронежский государственный университет, кафедра теории функций и геометрии, аспирант

Россия, 394018, г. Воронеж, Университетская пл., 1

E-mail: nickkolok@mail.ru, avdeev@math.vsu.ru

Avdeev Nikolai Nikolaevich

Voronezh State University, Department of Function Theory and Geometry, postgraduate student

Russia, 394018, Voronezh, University sq., 1

Зволинский Роман Евгеньевич

Воронежский государственный университет, кафедра теории функций и геометрии, аспирант

Россия, 394018, г. Воронеж, Университетская пл., 1

E-mail: roman.zvolinskiy@gmail.com , roman-zvolinskiy@mail.ru

Zvolinskii Roman Evgen'evich

Voronezh State University, Department of Function Theory and Geometry, postgraduate student

Russia, 394018, Voronezh, University sq., 1

Момот Екатерина Александровна

Воронежский государственный университет, кафедра функционального анализа и операторных уравнений, выпускница

Россия, 394018, г. Воронеж, Университетская пл., 1

Momot Ekaterina Alexandrovna

Voronezh State University, Department of Functional Analysis and Operator Equations (graduated from)

Russia, 394018, Voronezh, University sq., 1

\makeatletter
\ltx@iffilelater{biblatex-gost.def}{2017/02/01}%
{\toggletrue{bbx:gostbibliography}%
\renewcommand*{\revsdnamepunct}{\addcomma}}{}
\makeatother

\printbibliography
\printbibitembibliography
%\bibliography{literature}

%\begin{thebibliography}{99}
%\end{thebibliography}

\end{document}